---
title: "Final Project"
author: "Maisy Song"
date: "12/12/2021"
output: pdf_document
---

##Final Project

```{r library}
library(ggplot2)
library(faraway)
library(ISLR)
library(MASS)
```

```{r fileupload}
directory = ("Documents/uiuc/finalproject2/")
bikesdata = read.csv("bike_clean.csv")
```

## Point 1
Fit a saturated model (first model) including all first order terms for your predictors.
  - Pick one quantitative predictor, and interpret its fitted coefficient value.
  - Pick one categorical predictor, and interpret its fitted coefficient(s).  Be sure to specify the baseline level.
  
```{r}
bikes = subset(bikesdata, select = -c(X))
bikes = subset(bikes, select = c(Month, TemperatureC, Rented.Bike.Count, Hour, Humidity, Wind.speed.ms, Rainfallmm, Solar.Radiation..MJ.m2., Seasons))
# I have made some corrections/changes to variable names so printing out for some convenience
names(bikes)
```


```{r}
ggplot(bikes, mapping = aes(x = TemperatureC, y = Rented.Bike.Count)) + geom_point()
# round(cor(bikedata),4)
```

```{r}
firstmodel = lm(Rented.Bike.Count ~ ., bikes)
summary(firstmodel)
```
 
Here we can see how that there are some of the variables that aren't necessary/significant at level alpha = 0.05

```{r}
secondmodel = lm(Rented.Bike.Count ~ Hour + TemperatureC + Humidity + 
                   Solar.Radiation..MJ.m2. + Rainfallmm + Seasons, bikes)
summary(secondmodel)
```

```{r}
newbikes = subset(bikes, select = c(Rented.Bike.Count, Hour, TemperatureC, Humidity, Solar.Radiation..MJ.m2., Rainfallmm, Seasons))

second_order_model = lm(Rented.Bike.Count ~ .^2, newbikes)
# summary(second_order_model)

temp_model = lm(Rented.Bike.Count ~ . + Hour * TemperatureC, newbikes) # adj R^2 = 0.6135
# summary(temp_model)

test_model = lm(Rented.Bike.Count ~ . + TemperatureC * Humidity, newbikes)
# summary(test_model)

test_model2 = lm(Rented.Bike.Count ~ . + TemperatureC * Humidity + Hour * TemperatureC, newbikes)
summary(test_model2)
```


```{r}
# anova(lm(Rented.Bike.Count ~ 1, bikes), secondmodel)
# anova(secondmodel, firstmodel)


# # one interaction term added:
# anova(secondmodel, test_model)
# # two interaction terms added:
anova(secondmodel, test_model2)
# # we can see that the two interaction terms added model is more significant
```

```{r}
aic_backwards = step(test_model2, direction = 'backward')
finalmodel = lm(Rented.Bike.Count ~ Hour + TemperatureC + Humidity + Solar.Radiation..MJ.m2. + Rainfallmm + Seasons + TemperatureC * Humidity + Hour * TemperatureC, newbikes)
summary(finalmodel)
```

Here we can see that my final model is:
Rented.Bike.Count ~ Hour + TemperatureC + Humidity + Solar.Radiation..MJ.m2. + Rainfallmm + Seasons + TemperatureC * Humidity + Hour * TemperatureC

## Collinearity
```{r}
# here we want to omit categorical variable
firstmodel_nocat = lm(Rented.Bike.Count ~ . - Seasons, bikes)
finalmodel_nocat = lm(Rented.Bike.Count ~ Hour + TemperatureC + Humidity + Solar.Radiation..MJ.m2. + 
    Rainfallmm + TemperatureC * Humidity + Hour * TemperatureC, newbikes)
vif(firstmodel_nocat)
vif(finalmodel_nocat)
```
From the first model, I am not necessarily concerned about collinearity since all variables have vif values less than 10.
However, I can see how I can be concerned about collinearity of TemperatureC:Humidity and Hour:Temperature.

```{r}
summary(firstmodel)$adj.r.squared
summary(finalmodel)$adj.r.squared
```


## Unusual Points
```{r}
# plot(firstmodel)
plot(finalmodel)
```

## Boxcox
```{r}
boxcox(finalmodel, plotit = TRUE, lambda = seq(0, 0.5, 0.1))

transformed_mod = lm(((Rented.Bike.Count ^0.4 - 1)/0.4) ~ Hour + TemperatureC + Humidity + Solar.Radiation..MJ.m2. + Rainfallmm + Seasons + TemperatureC * Humidity + Hour * TemperatureC, newbikes)

plot(transformed_mod)
summary(transformed_mod)
```

## Statistical Test
```{r}
summary(finalmodel)
summary(transformed_mod)
```

```{r}
summary(firstmodel)$adj.r.squared
summary(transformed_mod)$adj.r.squared
```

```{r}
trans_nocat = lm(((Rented.Bike.Count ^0.4 - 1)/0.4) ~ Hour + TemperatureC + Humidity + Solar.Radiation..MJ.m2. + Rainfallmm + TemperatureC * Humidity + Hour * TemperatureC, newbikes)
vif(trans_nocat)
```

##RMSE
```{r}
sqrt(mean(resid(firstmodel) ^ 2))
sqrt(mean(resid(finalmodel) ^ 2))
sqrt(mean(resid(transformed_mod) ^ 2))
```

```{r}
# ggplot(bikes, mapping = aes(x = Solar.Radiation..MJ.m2., y = Rented.Bike.Count)) + geom_point()

mod2 = lm(((Rented.Bike.Count ^0.4 - 1)/0.4) ~ Hour + poly(TemperatureC,4) + Humidity + Solar.Radiation..MJ.m2. + Rainfallmm + Seasons + TemperatureC * Humidity + Hour * TemperatureC, newbikes)
# summary(mod2)
# plot(mod2)
# anova(transformed_mod, mod2)
# aic_backwards = step(mod2, direction = 'backward')
mod2_nocat = lm(((Rented.Bike.Count ^0.4 - 1)/0.4) ~ Hour + poly(TemperatureC,4) + Humidity + Solar.Radiation..MJ.m2. + Rainfallmm + TemperatureC * Humidity + Hour * TemperatureC, newbikes)
```


## Try fitting one of the values
```{r}
transformed_mod$coefficients
point1 = data.frame(Hour = 15, TemperatureC = -2.1, Humidity = 20, Solar.Radiation..MJ.m2. = 1.87, Rainfallmm = 0, Seasons = "Winter")

y = 33.492523097 + 0.384363947*point1$Hour + 0.408831369*point1$TemperatureC + -0.158968974*point1$Humidity + -0.937375254*point1$Solar.Radiation..MJ.m2.+ -1.308151968*point1$Rainfallmm + -0.002862839*point1$Humidity*point1$TemperatureC + 0.013283877*point1$Hour*point1$TemperatureC -11.082664910
((y*0.4) + 1)**(1/0.4)
# (Rented.Bike.Count ^0.4 - 1)/0.4)
vif(trans_nocat)
```







